---
title: Investigate exponential distribution in R and compare it with the Central Limit
  Theorem
author: "Thimmaraju Rudrappa"
date: "July 13, 2019"
output:
  html_document: default
  pdf_document: default
---

## Overview
The objective is to investigate the exponential distribution
and compare it to the Central Limit Theorem. 


## Simulations
Set variables lambda, exponentials, and seed.
```{r}
ECHO=TRUE
set.seed(1337)
lambda = 0.2
exponentials = 40
```

Run Simulations 
```{r}
sMeans = NULL
for (i in 1 : 1000) sMeans = c(sMeans, mean(rexp(exponentials, lambda)))
```


## Sample Mean Vs Theoretical Mean
#### Sample Mean
Calculate mean from the simulations with give the sample mean.
```{r}
mean(sMeans)
```

#### Theoretical Mean
Theoretical mean of an exponential distribution is lambda^-1.
```{r}
lambda^-1
```

#### Comparison
As we can see there is only a slight difference between the sample mean and the theoretical mean.
```{r}
abs(mean(sMeans)-lambda^-1)
```


## Sample Variance Vs Theoretical Variance
#### Sample Variance
Calculate the variance from the simulation means with give the sample
variance.
```{r}
var(sMeans)
```

#### Theoretical Variance
The theoretical variance of an exponential distribution is 
(lambda * sqrt(n))^-2.
```{r}
(lambda * sqrt(exponentials))^-2
```

#### Comparison
As we can see there is only a slight difference between the sample variance and theoretical variance.
```{r}
abs(var(sMeans)-(lambda * sqrt(exponentials))^-2)
```


## Distribution
Density histogram of the 1000 simulations. There is an overlay with a 
normal distribution that has a mean of lambda^-1 and standard deviation of 
(lambda*sqrt(n))^-1, the theoretical normal distribution for the simulations.
```{r}
library(ggplot2)
ggplot(data.frame(y=sMeans), aes(x=y)) + 
  geom_histogram(aes(y=..density..), binwidth=0.2, fill="#0072B2", 
                 color="black") +
  stat_function(fun=dnorm, arg=list(mean=lambda^-1, 
                                    sd=(lambda*sqrt(exponentials))^-1), 
                size=2) +
  labs(title="Plot of the Simulations", x="Simulation Mean")
```