---
title: "Effect of Vitamin C on Tooth Growth in Guinea Pigs"
author: "Thimmaraju Rudrappa"
date: "July 13, 2019"
output: html_document
---

### Overview
Objective of this analysis is to analyze the effect og Vitamin C on tooth growth in guinea pigs by using ToothGrowth dataset.

### Load the data and perform exploratory data analyses
```{r results='hide'}
library(datasets)
data(ToothGrowth)
str(ToothGrowth)
head(ToothGrowth)
summary(ToothGrowth)
```

```{r}
library(ggplot2)
t = ToothGrowth
levels(t$supp) <- c("Orange Juice", "Ascorbic Acid")
ggplot(t, aes(x=factor(dose), y=len)) + 
  facet_grid(.~supp) +
  geom_boxplot(aes(fill = supp), show_guide = FALSE) +
  labs(title="Guinea pig tooth length by dosage for each type of supplement", 
    x="Dose (mg/day)",
    y="Tooth Length")
```

### Basic summary
The plot shows increased vitamin C increases tooth growth and orange juice is more effective ascorbic acid.

### confidence intervals & hypothesis tests to compare tooth growth by supplement and dose
#### Hypothesis #1
Orange juice & ascorbic acid deliver the same tooth growth across the data set.
```{r}
hypoth1<-t.test(len ~ supp, data = t)
hypoth1$conf.int
hypoth1$p.value
```
The confidence intervals includes 0 and the p-value is greater than the 
threshold of 0.05.  The null hypothesis cannot be rejected.

#### Hypothesis #2
For the dosage of 0.5 mg/day, the two supplements caused the same tooth growth.
```{r}
hypoth2<-t.test(len ~ supp, data = subset(t, dose == 0.5))
hypoth2$conf.int
hypoth2$p.value
```
The confidence interval does not include 0 and the p-value is below the 0.05 
threshold. The null hypothesis can be rejected. The alternative hypothesis 
that 0.5 mg/day dosage of orange juice delivers more tooth growth than ascorbic 
acid is accepted.

#### Hypothesis #3
For the dosage of 1 mg/day, the two supplements deliver the same tooth growth
```{r}
hypoth3<-t.test(len ~ supp, data = subset(t, dose == 1))
hypoth3$conf.int
hypoth3$p.value
```
The confidence interval does not include 0 and the p-value is smaller than the 
0.05 threshold. The null hypothesis can be rejected. The alternative hypothesis 
that 1 mg/day dosage of orange juice delivers more tooth growth than ascorbic 
acid is accepted.

#### Hypothesis #4
For the dosage of 2 mg/day, the two supplements deliver the same tooth growth
```{r}
hypoth4<-t.test(len ~ supp, data = subset(t, dose == 2))
hypoth4$conf.int
hypoth4$p.value
```
The confidence interval does include 0 and the p-value is larger than the 
0.05 threshold. The null hypothesis cannot be rejected.


### Conclusions & assumptions
Orange juice delivers more tooth growth than ascorbic acid for dosages 0.5 & 
1.0. Orange juice and ascorbic acid deliver the same amount of tooth growth for 
dose amount 2.0 mg/day.  For the entire data set we cannot conclude orange juice
is more effective that ascorbic acid.

Assumptions

* Normal distribution of the tooth lengths
* No other unmeasured factors are affecting tooth length